---
title: "FY `r params$year` Giving Profile"
subtitle: "School of `r params$school` Alumni"
date: "`r format(Sys.Date(),'%d %B %Y')`"
output: 
  html_document:
    self_contained: TRUE
    keep_md: TRUE
    toc: TRUE
    toc_float: TRUE
    css: report.css
params:
  school: "Engineering"
  year: 2020
  geoType: "CBSA"
---

```{r setup, include=FALSE}
library(tidyverse)
# Currently in development
library(gt)

library(here)
library(keyring)

library(reticulate)

#### This establishes the connection to a SQL database. 

# Instead of displaying the connection information in free text: 
# connect(host = "my.server", user = "usr","password = "12345")
# Instead, it's wrapped in a custom function. The password itself 
# is stored in a keyring, and is *never* displayed.

con <- advancementtools::cx_Oracle(usr = "BI",
                                   pw = key_get("advance","BI","adv"))



knitr::opts_chunk$set(echo = FALSE,
                      message = F,
                      warning = F,
                      fig.showtext = TRUE,
                      # dpi = 72,
                      # The SQL connection needs to be passed to any SQL chunks,
                      # and can be set here.
                      connection = "con")
```

<!-- HTML documents can include HTML code in the body of the document, and will render appropriately 
     In this case, we're adding a logo to the header above the table of contents. 
-->

<script>
  /* Add Logo Header using JQuery*/
  $(document).ready(function() {
    $('#TOC').parent().prepend('<div id=\"nav_logo\"><img src=\"https://brand.virginia.edu/sites/default/files/styles/embedded_image_medium_2x/public/2017-08/rev_uva_logo_color.png\"></div>');
  });
</script>

## Overview

```{r loadData}
# Get the Biographic details
bio <- read_csv(here("Data/donorBio.csv")) %>% 
  # Only Alumni have a SCHOOL, replace missing values
  mutate(SCHOOL = replace_na(SCHOOL,"Non-Alumni"),
         # Format the Name for Display
         schoolName = gsub("Arts and Sciences","A&S",SCHOOL),
         # Create a variable for highlighting later on.
         highlight = factor(if_else(SCHOOL == params$school,"Y","N"),
                            levels = c("N","Y")))

# Read in the Giving information
giving <- read_csv(here("Data/newYrGiving.csv")) %>%
  # parameters can be accessed by `params${name}`
  filter(year == params$year)

# Join the biographic and giving data together
data <- bio %>% 
  inner_join(giving, by = "ID")

```

`r scales::comma(data %>% filter(SCHOOL == params$school) %>% nrow())` Alumni from the School of `r params$school` gave in the `r params$year` Fiscal Year. This accounts for `r scales::percent((data %>% filter(SCHOOL == params$school) %>% nrow())/nrow(data))` of donor participation.


```{r customTheme}
# Rather than copy and past all of the Theme elements for a ggplot2 element,
# save them as a pseudo-function.

customTheme <- theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.margin=unit(c(1,1,1,1),"cm"),
        axis.text = element_text(size = 20),
        axis.line.x = element_line(size = 1),
        axis.line.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(size = 0.5,
                                          colour = "darkgrey"),
        plot.background = element_rect(fill = "transparent",
                                      color = "transparent"))

```

```{r basicPlots, fig.width=12}
countPlot <- data %>% 
  ggplot(aes(x = schoolName, fill = highlight)) +
  geom_histogram(stat = "count") +
  coord_flip() +
  customTheme +
  labs(x=NULL, y=NULL,
       title = "Donor Participation") +
  scale_y_continuous(labels = scales::comma) +
  geom_hline(yintercept = 0,size = 1, color = "black") +
  scale_fill_manual(values = c("#232D4B","#E57200"))

dollars <- data %>%
  group_by(schoolName,highlight) %>% 
  summarize(total = sum(giving))

dollarsPlot <-  dollars %>% 
  ggplot(aes(x = schoolName, y = total, fill = highlight)) +
  geom_histogram(stat="identity") + 
  labs(x=NULL, y=NULL,
       title = "Dollars Raised (log scale)") +
  coord_flip() +
  customTheme + 
  theme(axis.text.y = element_blank(),
        axis.line.x = element_line(size = 1)) +
  geom_hline(yintercept = 1,size = 1, color = "black") +
  scale_y_log10(breaks = 10^(0:9),
    labels = c("$0","$10","$100",
               "$1K","$10K","$100K",
               "$1M","$10M","$100M","$1T")) +
  geom_text(aes(label = scales::dollar(total),
                color = highlight),
            size = 5, hjust= 1.2) +
  scale_fill_manual(values = c("#232D4B","#E57200")) +
  scale_color_manual(values = c("white","#232D4B"))

gridExtra::grid.arrange(countPlot,dollarsPlot,ncol = 2)
```


## By Region

```{r geoPrep}
geoType <- params$geoType

# While not used here, the `glue` package can properly format multiple arguments
# for injection into a SQL chunk.
geo <- data %>% 
  filter(!is.na(ZIPCODE)) %>% 
  distinct() %>% 
  pull(ZIPCODE)

geo <- glue::glue_sql("{geo*}",.con = con)
```

Of the `r scales::comma(nrow(data))` donors in FY `r params$year`, `r scales::comma(nrow(data %>% filter(!is.na(ZIPCODE))))` (`r scales::percent(nrow(data %>% filter(!is.na(ZIPCODE)))/nrow(data))`) have a valid Zip Code.

```{sql, output.var = "cx_geography"}
-- This SQL query creates a crosswalk between Zip codes and Core-Based 
-- Statistical Areas (CBSA), Metropolitan Statistical Areas (MSA) and Combined 
-- Statistical Areas (CSAs) via the 5-digit county-level FIPS code.

-- Note the "output.var" chunk argument, which saves the results as an R object.

WITH geo as (
    SELECT  FIP5
            ,title
            ,geography
    FROM    cx_geo
    UNPIVOT (title for geography in (
                                    county as 'County'
                                    ,cbsa_title     as 'CBSA'
                                    ,csa_title as 'CSA'
                                    ,msa_title as 'MSA'
                                    )
            )
    )


SELECT      DISTINCT ZIP            ZIPCODE
            ,title
FROM        cx_zip
INNER JOIN  geo
USING       (FIP5)
-- SQL chunks can use R variables using `?{variable_name}`
WHERE       geography = ?geoType
```

```{r crosswalk}
# Combine the data with the crosswalks obtained by SQL
combo <- data %>% 
  left_join(cx_geography, by = "ZIPCODE") %>% 
  # A Zip code may have multiple matches, this creates one row per Zip.
  group_by(ZIPCODE) %>% 
  slice(1) %>%
  ungroup()


topN <- 10
```

`r scales::comma(nrow(combo %>% filter(is.na(TITLE))))` entities reside outside of a `r geoType`. The table below shows the top `r topN` regions by count.

```{r topGeo}
comboTable <- combo %>% 
  filter(!is.na(TITLE)) %>% 
  count(TITLE,highlight, sort = T) %>% 
  spread(highlight, n, fill = 0)

# Error Handling for when No alumni participated
if (!"Y" %in% names(comboTable)){
  comboTable$Y = 0
}

comboTable <- comboTable %>% 
  mutate(total = Y+N,
         per = Y/total) %>% 
  arrange(desc(Y),total) %>% 
  slice(1:topN)

comboTable %>% 
  gt() %>% 
  cols_hide(columns = vars(N)) %>% 
  cols_label(TITLE = params$geoType,
             Y = "Alumni",
             total = "Total Donors",
             per = "Alumni %") %>% 
  fmt_percent(columns = vars(per),
              decimals = 0) %>% 
  tab_header(title = md(sprintf("School of **%s** Alumni",params$school)),
             subtitle = sprintf("Donor Participation in FY %i",params$year))
  
  
```

**`r comboTable$TITLE[1]`** was the `r params$geoType` with the most number of `r params$school` alumni donors (`r comboTable$Y[1]`).

```{python geoHist}
import pandas as pd
import numpy as np

np.random.seed(2020)

titles = r.combo['TITLE'].value_counts()

# Remove NA values
t2 = titles.drop('NA')

pl = t2.plot.hist(figsize = (14,8),
                  title = "%s Histogram" % (r.geoType),
                  color = '#232D4B')

pl.set_xlabel("Entites per %s" % (r.geoType))

# pl.show()
```

`r scales::comma(py$titles["NA"])` individuals do __not__ have a valid `r geoType`.

```{r geoTable}
geoTable <- table(py$titles)


maxRegions <- names(geoTable[geoTable == max(geoTable)])
```

`r scales::percent(sum(geoTable[geoTable == max(geoTable)])/sum(geoTable))` of regions have `r paste(names(geoTable[geoTable == max(geoTable)]),collapse = " or ")` `r if(maxRegions == 1){"entity"}else{"entities"}`

## Session Info

Analysis and Report generated on `r sessionInfo()[["running"]]` using `r sessionInfo()[["R.version"]]$version.string`.

```{r packageList, results='asis'}
# Results = 'asis' renders the output as if it were text in the markdown doc. In
# this case, the custom function (based on the `packageList` chunk in "The
# Technical" Rmd document) renders as a bulletted list.

# Lists attached packages
advancementtools::sessionStats()
```

<!-- Insert Footnotes Below -->
```{r disconnnect,results='hide'}
advancementtools::dx_Oracle(con)
```


```{r render, eval = F}
rmarkdown::render(here("Markdown/3-TheShowstopper.Rmd"),
                  output_dir = here("Reports"),
                  quiet = T)
```

